Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} Model System

### 2014-05-29 trunk/paper_enahpc14/Sec4:Experimental-results.tex

Averaged TTS and ETS results  over 10 repetitions of 24-hours simulation on
\tinto are shown in  Figure~\ref{fig:4}. The corresponding results for
the  averaged total  power are  also shown  in  Table~\ref{tab:1}.  As
expected, the average total power consumption by \cosmoart while using
the \emph{degraded} MPI policy is about  2\,\% less than that consumed by the
\emph{aggressive} one. This is  due to  the reduction of CPU usage (of about 50\,\%) while using  the \emph{degraded} MPI policy, in
contrast with the \emph{aggressive} policy that always yields to 100\,\% of CPU utilization while waiting for an incoming message. From the point of  view of the execution time, there is a
slight reduction  for the \emph{degraded} MPI  policy. One can  expect that the
costs of the context switches produced in the \emph{degraded} MPI policy can lead to
an  increase  of  the   execution  time,  however,  applications  with
extensive computing,  as e.g.  \cosmoart, may show  better performance
with the \emph{degraded} policy. This effect works in  favor with our results
since the energy consumed is also reduced (2\,\%) just by changing the
behavior of the OpenMPI library  to the \emph{degraded} policy, i.e, with the
\texttt{--mca    mpi\_yield\_when\_idle 1}   parameter   set    in   the
\texttt{mpirun} command.

\begin{figure}[htbf]
  \centering
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-0.eps}
  \caption{Time-to-solution and  energy-to-solution comparison between
    \pilat and \monch for a 24-hours simulation.}
  \label{fig:3}
\end{figure}

\begin{figure}[htbf]
  \centering
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-1.eps}
  \vspace{-1cm}
  \caption{Mean time-to-solution and  energy-to-solution on \tinto for
    a 24-hours simulation using the  \emph{aggressive} and \emph{degraded} MPI policies. The
    95\,\%  confidence  intervals of  Student's  $t$ distribution  (10
    samples) clearly  indicates an improvement for the \emph{degraded} policy in
    both TTS and ETS.}
  \label{fig:4}
\end{figure}

\begin{figure*}[ht]
  \centering
  \hspace{0.8cm} \scalebox{0.62}{\input{Figs/24hour.pstex_t}}
  \caption{24-hours simulation power  trace  using  the \emph{aggressive}  and
    \emph{degraded} MPI policies on \tinto.}
  \label{fig:5}
\end{figure*}

@@

In order to obtain an initial overview of the power consumption of the
system  model,  we  obtained  power  profiles  leveraging  our  \pmlib
power-measurement    framework   over    a    1-day   simulation    of
\cosmoart. These profiles, produced with both \emph{aggressive} and \emph{degraded} MPI
policies,  are depicted in  Figure~\ref{fig:5}. The  first observation
for the \emph{aggressive} MPI policy is that the first and last hours of the day
consume slightly less  power than the midday hours.  We relate these
variations  due   to  the  different   chemical  reactions,  aerosols,
radiations  and clouds  formation processes,  computed by  ART, taking
place during  daylight hours  but not during  night. In this  case the
peak power consumption is about  3821\,W. Looking at the power profile
using the \emph{degraded}  MPI policy, we observe that  the 1-hour pattern is
repeated along  the 24-hours, favoring  the reduction of  power during
night  hours in  a  more noticeable  way  than that  with the   \emph{aggressive}
policy.   The peak  power consumption  during day  hours  while the
 \emph{degraded}  MPI policy  is leveraged  is  about 3768\,W,  which means  a
reduction of around  50\,W with respect to the  \emph{aggressive} mode. Also, we
notice systematic  power drops for each simulated  hour.  We associate
them  to periods  in  which  the processes  were waiting  for
incoming messages  or synchronizations, thus  yielding the CPU  to a lower
usage. Our final observation is the reduction of the TTS by 2\,\% with
respect to the  \emph{degraded} MPI policy.

>>>>

Averaged TTS and ETS results  over 10 repetitions of 24-hour simulation on
\tinto are shown in  Figure~\ref{fig:4}. The corresponding results for
the  averaged total  power are  also shown  in  Table~\ref{tab:1}.  As
expected, the average total power consumption by \cosmoart while using
the \emph{degraded} MPI policy is about  2\,\% less than that consumed by the
\emph{aggressive} one. This is  due to  the reduction of CPU usage (of about 50\,\%) while using  the \emph{degraded} MPI policy, in
contrast with the \emph{aggressive} policy that always yields to 100\,\% of CPU utilization while waiting for an incoming message. From the point of  view of the execution time, there is a
slight reduction  for the \emph{degraded} MPI  policy. One can  expect that the
costs of the context switches produced in the \emph{degraded} MPI policy can lead to
an  increase  of  the   execution  time,  however,  applications  with
extensive computing,  as e.g.  \cosmoart, may show  better performance
with the \emph{degraded} policy. This effect works in  favor with our results
since the energy consumed is also reduced (2\,\%) just by changing the
behavior of the OpenMPI library  to the \emph{degraded} policy, i.e, with the
\texttt{--mca    mpi\_yield\_when\_idle 1}   parameter   set    in   the
\texttt{mpirun} command.

\begin{figure}[htbf]
  \centering
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-0.eps}
  \caption{Time-to-solution and  energy-to-solution comparison between
    \pilat and \monch for a 24-hour simulation.}
  \label{fig:3}
\end{figure}

\begin{figure}[htbf]
  \centering
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-1.eps}
  \vspace{-1cm}
  \caption{Mean time-to-solution and  energy-to-solution on \tinto for
    a 24-hour simulation using the  \emph{aggressive} and \emph{degraded} MPI policies. The
    95\,\%  confidence  intervals of  Student's  $t$ distribution  (10
    samples) clearly  indicates an improvement for the \emph{degraded} policy in
    both TTS and ETS.}
  \label{fig:4}
\end{figure}

\begin{figure*}[ht]
  \centering
  \hspace{0.8cm} \scalebox{0.62}{\input{Figs/24hour.pstex_t}}
  \caption{24-hour simulation power  trace  using  the \emph{aggressive}  and
    \emph{degraded} MPI policies on \tinto.}
  \label{fig:5}
\end{figure*}

@@

In order to obtain an initial overview of the power consumption of the
system  model,  we  obtained  power  profiles  leveraging  our  \pmlib
power-measurement    framework   over    a    1-day   simulation    of
\cosmoart. These profiles, produced with both \emph{aggressive} and \emph{degraded} MPI
policies,  are depicted in  Figure~\ref{fig:5}. The  first observation
for the \emph{aggressive} MPI policy is that the first and last hours of the day
consume slightly less  power than the midday hours.  We relate these
variations  due   to  the  different   chemical  reactions,  aerosols,
radiations  and clouds  formation processes,  computed by  ART, taking
place during  daylight hours  but not during  night. In this  case the
peak power consumption is about  3821\,W. Looking at the power profile
using the \emph{degraded}  MPI policy, we observe that  the 1-hour pattern is
repeated along  the 24 hours, favoring  the reduction of  power during
night  hours in  a  more noticeable  way  than that  with the   \emph{aggressive}
policy.   The peak  power consumption  during day  hours  while the
 \emph{degraded}  MPI policy  is leveraged  is  about 3768\,W,  which means  a
reduction of around  50\,W with respect to the  \emph{aggressive} mode. Also, we
notice systematic  power drops for each simulated  hour.  We associate
them  to periods  in  which  the processes  were waiting  for
incoming messages  or synchronizations, thus  yielding the CPU  to a lower
usage. Our final observation is the reduction of the TTS by 2\,\% with
respect to the  \emph{degraded} MPI policy.
